Supernova Host Galaxies in the Dark Energy Survey: III. Constraining the SN Ia delay time dstribution

### 2021-01-04 draft.tex

Another method of inferring progenitor channels is the study of the SN Ia delay time distribution (DTD). The DTD can be interpreted as the rate at which SNe Ia occur as a function of the delay time $\tau$ since an episode of star-formation, and thus carries characteristic signatures of the progenitor channel (see \citealt{Wang2012} for a review). While the SD scenario is able to produce a broad range of functional forms of the DTD, the DD scenario predicts a power law of the form $t^{\beta}$, with $\beta \sim -1$ \citep[e.g][]{Ruiter2009,Mennekens2010}.

@@

@@

\section{Data \label{sec:data}}
\subsection{Dark Energy Survey supernova programme \label{subsec:des}}
The DES Supernova Programme (DES-SN) is a transient survey based on five six-month seasons of observations with the Dark Energy Camera (DECam; \citealt{Flaugher2015}), designed primarily to measure the light curves of SNe Ia for use as cosmological distance indicators. Transients were detected and processed using a difference imaging program \citep{Kessler2015} and vetted using a machine learning (ML) algorithm \citep{Goldstein2015}. Spectral follow-up was performed on a suite of large optical telescopes \citet{Smith2020b}, leading to the initial publication of a measurement of cosmological parameters using 207 spectroscopically confirmed SNe (\citealt{DESCollaboration2018a}, and references therein). 
\subsection{Supernovae \label{subsec:host_sample}}
\subsubsection{Photometric classification and quality cuts \label{subsubsec:sn_classify}}
With over 30,000 transients passing ML cuts it was not possible to obtain spectral follow up of every object. Instead we make use of photometric classification techniques, in particular the recurrent neural network \texttt{superNNova} (SNN; \citealt{Moller2019}). We derive our SN Ia sample following the approach of \citet{Scolnic2020}, but use a threshold probability of $P(\mathrm{Ia})\geq0.5)$. The photometricially classified SNe Ia are then passed through the \texttt{SALT2} framework \citep{Betoule2014} and are subject to quality cuts on the light curve width (stretch; $x_1$) and colour ($c$) parameters: $-3 \leq x_1 \leq 3$, $-0.3 \leq c \leq 0.3$. These cuts help to reduce the potential contamination from core-collapse SNe (CCSNe) and outlying SNe Ia whose physical mechanisms may diverge from the general cosmologically useful population.

@@

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/rate_vs_mass_DTD_fit_beta_norm.png}
    \caption{The rate per galaxy of SNe Ia as a function of stellar mass. The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citet{Childress2014} model}
    \label{fig:rate_fit}
\end{figure}
\section{Incompleteness corrections \label{sec:incompleteness}}

Although Eq. \ref{eq:rate1} serves as an approximation to the SN rate per galaxy per year, both the SN and field galaxy samples introduced in Sections \ref{subsec:host_sample} and \ref{subsec:field_sample} are affected by incompleteness, which is likely to be the dominant systematic effect in the analysis. In this section we describe our method of correcting for various sources incompleteness in the data.

@@

\section{Modeling the rate of type Ia supernovae \label{sec:method}}

We represent the SN Ia DTD by a power law of the form:

\begin{equation}
    \Phi(\tau) = A\left(\frac{\tau}{\mathrm{Gyr}}\right)^{\beta}\,,
\end{equation} 
where $\tau$ is the time since a burst of star formation, and $A$ is a normalisation at $\tau=1\mathrm{Gyr}$. If all stars in a galaxy were formed at a single epoch $t_f$, the DTD would describe the rate of SNe at an observation time $\tau = t_0$. However in reality galaxies are formed by the gradual build up of stellar mass over several epochs of star formation -- the distribution of this mass build up is known as the star-formation history (SFH). The rate of SNe Ia is thus the sum of the DTD evaluated for each epoch of star formation, multiplied by the stellar mass formed in that epoch. Mathematically this is represented by the convolution of the DTD and SFH:

\begin{equation}
    R_{\mathrm{G}} = \int_{t_0}^{\infty} \psi(t_0-\tau)\Phi(\tau)\,d\tau \,,
\end{equation}
where $t_0$ is the epoch at which the galaxy is observed, and $\psi$ is the SFH.

\subsection{Modeling the star formation histories of galaxies \label{subsec:method_sfh}}
To model the SFH of galaxies as a function of their stellar mass we adopt the model of \citet{Childress2014}. In the model, galaxies are expected to evolve smoothly along the so-called "main sequence of star formation" whereby the SFR of a galaxy of stellar mass $M_*$ at redshift $z$ is determined by a simple relationship (the SMz relation, \citealt{Zahid2012}). Specifically, we implement their "fine-tuned" model (C14 Eq. A7), whereby the SMz flattens above $z\sim2$ in line with observations \citep{Stark2013}. We adjust this model slightly and arrive at a final SMz:

@@

The result of our toy model is shown in Fig. \ref{fig:SFHs}. Galaxies with high masses at $z=0$ formed the vast majority of their stars in the first few Gyr and are now likely to be passive with old stellar populations, while low-mass galaxies are currently strongly star-forming with a vast majority of young stars. Galaxies with masses around $10^10 \mathrm{M}_{\odot}$ are composed of a mixture of young and older stellar populations.

\section{Results \label{sec:rates}}

The rate of SNe per galaxy ($R_{\mathrm{G}}$) per year in a transient survey can be approximated by the simple equation:
\begin{equation}
    R_{\mathrm{G}} = \frac{N_{\mathrm{SN}}}{N_{\mathrm{G}}} \frac{A_{\mathrm{G}}}{A_{\mathrm{SN}}} \frac{1}{T_{\mathrm{SN  survey}}}\,,
\label{eq:rate1}
\end{equation}
where $N_{\mathrm{SN}}$ and $N_{\mathrm{G}}$ are the respective number of SNe and galaxies detected, $A_{\mathrm{SN}}$ and $A_{\mathrm{G}}$ the sky areas from which the SNe and G samples were taken, and $T$ the duration of the SN survey in years. As described in Section \ref{sec:incompleteness}, the values of $N_{\mathrm{SN}}$ and $N_{\mathrm{G}}$ that we observe are underestimates of the true numbers due to observational incompleteness. We thus estimate the intrinsic numbers of SNe and galaxies by multiplying the observed numbers by their respective incompleteness corrections,

\begin{equation}
    N_{\mathrm{SN,intrinsic}} = \eta_{\mathrm{SN}} \cdot N_{\mathrm{SN, obs}}
\end{equation}
and
\begin{equation}
    N_{\mathrm{G,intrinsic}} = \eta_{\mathrm{G}} \cdot N_{\mathrm{G, obs}}
\end{equation}

By binning both the SN hosts and field galaxies by their stellar mass we can approximate the function $R_{\mathrm{G}}(M_*)$.

@@

@@

\section{Conclusions}

The last numbered section should briefly summarise what has been done, and describe
the final conclusions which the authors draw from their work.

>>>>

Another method of inferring progenitor channels is the study of the SN Ia delay time distribution (DTD). The DTD can be interpreted as the rate at which SNe Ia occur as a function of the delay time $\tau$ since an episode of star-formation, and thus carries characteristic signatures of the progenitor channel (see \citealt{Wang2012} for a review). While the SD scenario is able to produce a broad range of functional forms of the DTD, the DD scenario predicts a power law of the form $\tau^{\beta}$, with $\beta \sim -1$ \citep[e.g][]{Ruiter2009,Mennekens2010}.

@@

We introduce the SN host and field galaxy samples in Section \ref{sec:data} and describe our handling of incompleteness in Section \ref{sec:incompleteness}. We outline the modelling of star-formation histories in Section \ref{sec:method}, and present our results in Section \ref{sec:rates}. In Section \ref{sec:discussion} we discuss the robustness of the result and its meaning in a wider context, before concluding in Section \ref{sec:conclusion}.
Where relevant, we assume a spatially flat $\Lambda$CDM cosmology with $\Omega_m = 0.3$ and $H_0 = 70 \mathrm{ km s}^{-1}\mathrm{Mpc}^{-1}$. Unless otherwise stated, we assume Gaussian uncertainties quoted at the $1\sigma$ level. Magnitudes are quoted in the AB system \citep{Oke1983}.

@@

%%%% DATA %%%%
\section{Data}
\label{sec:data}
\subsection{Dark Energy Survey supernova programme \label{subsec:des}}
The DES Supernova Programme (DES-SN) is a transient survey based on five six-month seasons of observations with the Dark Energy Camera (DECam; \citealt{Flaugher2015}), designed primarily to measure the light curves of SNe Ia for use as cosmological distance indicators. Transients were detected and processed using a difference imaging program \citep{Kessler2015} and vetted using a machine learning (ML) algorithm \citep{Goldstein2015}. Spectral follow-up was performed on a suite of large optical telescopes \citet{Smith2020b}, leading to the initial publication of a measurement of cosmological parameters using 207 spectroscopically confirmed SNe (\citealt{DESCollaboration2018a}, and references therein). 
\subsection{Supernovae \label{subsec:host_sample}}
\subsubsection{Photometric classification and quality cuts \label{subsubsec:sn_classify}}
With over 30,000 transients passing ML cuts it was not possible to obtain spectral follow up of every object. Instead we make use of photometric classification techniques, in particular the recurrent neural network \texttt{superNNova} (SNN; \citealt{Moller2019}). We derive our SN Ia sample following the approach of \citet{Scolnic2020}, but use a threshold probability of $P(\mathrm{Ia})\geq0.5)$. The photometricially classified SNe Ia are then passed through the \texttt{SALT2} framework \citep{Betoule2014} and are subject to quality cuts on the light curve width (stretch; $x_1$) and colour ($c$) parameters: $-3 \leq x_1 \leq 3$, $-0.3 \leq c \leq 0.3$. These cuts help to reduce the potential contamination from core-collapse SNe (CCSNe) and outlying SNe Ia whose physical mechanisms may diverge from the general cosmologically useful population.

@@

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/rate_vs_mass_all.png}
    \caption{The rate per galaxy of SNe Ia as a function of stellar mass.}%The blue lines indicate the best fitting model based on a $t^{\beta}$ power law DTD convolved with SFHs based on the \citet{Childress2014} model}
    \label{fig:rate_raw}
\end{figure}
%%%% INCOMPLETENESS %%%%
\section{Incompleteness corrections}
\label{sec:incompleteness}
Although Eq. \ref{eq:rate1} serves as an approximation to the SN rate per galaxy per year, both the SN and field galaxy samples introduced in Sections \ref{subsec:host_sample} and \ref{subsec:field_sample} are affected by incompleteness, which is likely to be the dominant systematic effect in the analysis. In this section we describe our method of correcting for various sources incompleteness in the data.

@@

%%% MODEL %%%
\section{Modelling the rate of type Ia supernovae}
 \label{sec:method}
We represent the SN Ia DTD by a power law of the form:
\begin{equation}
    \Phi(\tau) = A\left(\frac{\tau}{\mathrm{Gyr}}\right)^{\beta}\,,
\end{equation} 
where $\tau$ is the time since a burst of star formation, and $A$ is a normalisation at $\tau=1\mathrm{Gyr}$. If all stars in a galaxy were formed at a single epoch $t_f$, the DTD would describe the rate of SNe at an observation time $\tau = t_0$. However in reality galaxies are formed by the gradual build up of stellar mass over several epochs of star formation -- the distribution of this mass build up is known as the star-formation history (SFH). The rate of SNe Ia is thus the sum of the DTD evaluated for each epoch of star formation, multiplied by the stellar mass formed in that epoch. Mathematically this is represented by the convolution of the DTD and SFH:

\begin{equation}
    R_{\mathrm{G}} = \int_{t_0}^{\infty} \psi(t_0-\tau)\Phi(\tau)\,d\tau \,,
    \label{eq:galaxy_rate}
\end{equation}
where $t_0$ is the epoch at which the galaxy is observed, and $\psi$ is the SFH.

In previous work \citep[e.g.][]{Strolger2004,Maoz2012} it has been common to determine the SFH for every galaxy in the survey via SED fitting. Eq. \ref{eq:galaxy_rate} is then used to calculate an expected number of SNe in each galaxy given the effective survey time, which is compared to the observed number in that galaxy (usually 0 and occasionally 1) using Poisson statistics. This method relies on either photometry covering several wavelength bands, or optical spectra with high signal-to-noise ratio (S/N), in order to distinguish accurately between SFHs. Such accuracy is not possible for the DES sample: we do not possess spectra of all galaxies in the field, and in some cases only have the four optical bands available -- and have a maximum of eight from NUV to NIR -- from which to infer a SFH. This lack of detail is compounded by our relying on photometric redshifts which add an extra layer of uncertainty to the calculation.

Instead of determining individual SFHs for each galaxy in our field sample, we use an empirical approach to estimate mean SFHs for galaxies as a function of their stellar mass at any given redshift (or equivalently, time $t_0$), that can be represented as $\hat \psi \left(t_0 -\tau; M_* \right)$. This method paves the way for modelling the SN Ia rate as a function of stellar mass by combining the mean SFHs and the DTD through Eq. \ref{eq:galaxy_rate}, and thus placing constraints on the DTD.

\subsection{Modelling the star formation histories of galaxies \label{subsec:method_sfh}}

To model the SFH of galaxies as a function of their stellar mass we adopt the prescription of stellar mass assembly of \citet{Childress2014}, which draws on the work of \citet{Zahid2012}. In the model, galaxies are expected to evolve smoothly along the so-called "main sequence of star formation" whereby the SFR of a galaxy of stellar mass $M_*$ at redshift $z$ is determined by a simple relationship (the SMz relation, \citealt{Zahid2012}). Specifically, we implement their "fine-tuned" model (C14 Eq. A7), whereby the SMz flattens above $z\sim2$ in line with observations \citep{Stark2013}. We adjust this model slightly and arrive at a final SMz:

@@

The result of our toy model is shown in Fig. \ref{fig:SFHs}. Galaxies with high masses at $z=0$ formed the vast majority of their stars in the first few Gyr and are now likely to be passive with old stellar populations, while low-mass galaxies are currently strongly star-forming with a vast majority of young stars. Galaxies with masses around $10^{10} \mathrm{M}_{\odot}$ are composed of a mixture of young and older stellar populations.

%%%%%% RESULTS %%%%%%%%

\section{Results }
\label{sec:rates}
The rate of SNe per galaxy ($R_{\mathrm{G}}$) per year in a transient survey can be approximated by the equation:
\begin{equation}
    R_{\mathrm{G}} = \frac{N_{\mathrm{SN}}}{N_{\mathrm{G}}} \frac{V_{\mathrm{G}}}{V_{\mathrm{SN}}}\frac{1}{T_{\mathrm{SN  survey}}}\,,
\label{eq:rate1}
\end{equation}
where $N_{\mathrm{SN}}$ and $N_{\mathrm{G}}$ are the respective number of SNe and galaxies detected, $V_{\mathrm{SN}}$ and $V_{\mathrm{G}}$ the volumes from which the SNe and G samples were taken, and $T$ the duration of the SN survey in years.

As described in Section \ref{sec:incompleteness}, the values of $N_{\mathrm{SN}}$ and $N_{\mathrm{G}}$ that we observe are underestimates of the true numbers due to observational incompleteness -- we do not detect and count all SNe in the volume $V_{\mathrm{SN}}$, nor do we detect and count all galaxies in the volume $V_{\mathrm{G}}$. We thus estimate the intrinsic numbers of SNe and galaxies by multiplying the observed numbers by their respective incompleteness corrections calculated in Sections \ref{subsec:incompleteness_SNe} and \ref{subsec:incompletenss_SN_hosts} for SNe and Section \ref{subsec:incompleteness_field} for field galaxies:

\begin{equation}
    N_{\mathrm{SN,intrinsic}} = \sum_i^{n_{\mathrm{SN}}}\eta_{\mathrm{SN,}i}\,,
    \label{eq:corr_SN}
\end{equation}
and
\begin{equation}
    N_{\mathrm{G,intrinsic}} = \sum_j^{n_{\mathrm{G}}} \eta_{\mathrm{G,}j} 
    \label{eq:corr_G}
\end{equation}
for each SN $i$ and galaxy $j$.

@@

By binning both the SN hosts and field galaxies by their stellar mass we measure the mean SN Ia rate per galaxy as a function of stellar mass, $R_{\mathrm{G}}(M_*)$. We employ a bootstrap Monte-Carlo approach in order to estimate the uncertainty in the value of the rate in each stellar mass bin. The probability distribution function (PDF) of each galaxy's stellar mass is represented by the sum of two half Gaussian distributions to represent the asymmetric positive and negative uncertainties derived in the SED fitting stage (Section \ref{sec:data}). For each SN host and field galaxy, we take 100 samples of its stellar mass by drawing at random from its PDF. For each of the 100 samples we calculate the SN rate per galaxy in each stellar mass bin via Eq \ref{eq:rate1}, using the completeness-corrected values of $N_{\mathrm{SN}}$ and $N_{\mathrm{SN}}$ from Eqs. \ref{eq:corr_SN} and \ref{eq:corr_G} respectively.

The rate of SNe Ia per galaxy as a function of stellar mass is shown in Fig. \ref{fig:rate_raw}.

@@

\section{Discussion}
\label{sec:discussion}

\section{Conclusions}
\label{sec:conclusion}
The last numbered section should briefly summarise what has been done, and describe
the final conclusions which the authors draw from their work.
